---
title: "Asessment Rmd file"
author: "B209978"
date: "17/06/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **R code to extract relevant data**
*https://github.com/B209978/B209978_assessment/tree/master/AssessmentFiles*

## Loading packages including NHSR dataset

```{r Loading_datasets, message=FALSE, warning=FALSE}
# required packages are:
library(NHSRdatasets)
library(tidyverse)
library(knitr)
library(here)
library(scales)
library(caret)
library(dataMeta)
```

### Load and explore NHS England A&E attendance data
#### Examining structure and completeness of dataset
``` {r load_data, echo=FALSE}
# review dataset and display any missing values
data(ae_attendances)
ae<-ae_attendances
glimpse(ae)
ae
ae %>% 
  map(is.na) %>%
map(sum)
```

### Adding index for later linkage
``` {r index_and_save, echo=FALSE}
# add index column
ae <- rowid_to_column(ae, "index")
# display for confirmation
ae %>%
  # Set the period column to show in month-year format
  mutate_at(vars(period), format, "%b-%y") %>% 
  # Set the numeric columns to have a comma at the 1000's place
  mutate_at(vars(attendances, breaches, admissions), comma) %>%
  # Show the first 10 rows
  head(10) %>%
  # Format as a table
  kable()
# save file
write_csv(ae, here("RawData", "ae_attendances.csv"))
```

### Filter data
#### Removed all hospitals except consultant led Emergency Departments covered by selected ambulance service

``` {r, filter_data, echo=FALSE}
# filter to include certain hospitals
# filter to include type 1 hospitals
# filter to remove type column
ae<-ae %>% select(index, period, org_code, type, attendances, breaches, admissions)%>% 
  filter(org_code == "RKB"|org_code == "RRK"|org_code == "RXK"|org_code == "RLT"| org_code == "RJC"|
           org_code == "RWP"|org_code == "RLQ"|org_code == "RJE"| org_code == "RL4"|org_code == "RNA"|
           org_code == "RBK"|org_code == "RQW"|org_code == "RXW",type == 1) %>% 
  select(-type)

ae
```


### Calculate metrics

``` {r calculations, echo=FALSE}
# calculate and view breach performance
WM_performance <- ae %>%
  group_by(period) %>%
  summarise_at(vars(attendances, breaches), sum) %>%
  mutate(breach_performance = 1 - breaches / attendances)
# glimpse(WM_performance)

# calculate and view admission rate
WM_admission <- ae %>%
  group_by(period) %>%
  summarise_at(vars(attendances, admissions), sum) %>%
  mutate(admission_rate = admissions / attendances)
# glimpse(WM_admission)

# calculate and view regional attendances
WM_attendance <- ae %>%
  group_by(period) %>%
  summarise_at(vars(attendances), sum)
# glimpse(WM_attendance)

WM_metrics<-ae %>% 
  group_by(period) %>% 
  summarise_at(vars(attendances, breaches, admissions), sum) %>% 
  mutate(breach_performance = 1 - breaches / attendances,
         admission_rate = admissions / attendances)
glimpse(WM_metrics)

```

### Brief visualisation of regional patterns
``` {r graphs, echo=FALSE}
# plot performance
ggplot(WM_performance, aes(period, breach_performance)) +
  geom_line(color = "darkcyan") +
  geom_point(color = "darkcyan") +
  scale_y_continuous(labels = percent) +
  scale_x_date(date_labels = "%b-%y", date_breaks = "12 month")+
  labs(x = "Month of attendance",
       y = "% of A&E attendances that met the four hour standard",
       title = "Diminishing accident and emergency (A&E) performance",
       caption = "Source: NHSRdatasets")

# plot admission rate
ggplot(WM_admission, aes(period, admission_rate)) +
  geom_line(color = "darkcyan") +
  geom_point(color = "darkcyan") +
  scale_y_continuous(labels = percent) +
  scale_x_date(date_labels = "%b-%y", date_breaks = "12 month")+
  labs(x = "Month of attendance",
       y = "% of A&E attendances that result in admission",
       title = "Increasing accident and emergency (A&E) admission rate",
       caption = "Source: NHSRdatasets")

# plot regional admissions
ggplot(WM_attendance, aes(period, attendances)) +
  geom_line(color = "darkcyan") +
  geom_point(color = "darkcyan") +
  scale_y_continuous() +
  scale_x_date(date_labels = "%b-%y", date_breaks = "12 month")+
  labs(x = "Month of attendance",
       y = "Number of A&E attendances",
       title = "Increasing accident and emergency (A&E) attendances",
       caption = "Source: NHSRdatasets")
```

### Save data subset
``` {r save_subset, echo=FALSE}
# view subset
glimpse(ae)
# save subset
write_csv(ae, here("RawData", "ae_attendances_WM_4hr_perfom.csv"))
```

### Divide dataset into training, marker and test
``` {r divide_dataset, echo=FALSE}
# confirm length of dataset
# nrow(ae)


# calculate proportion needed for small training datset 
prop<-(1-(15/nrow(ae)))

# ensure consistant random number generator
set.seed(333)
# create training partition
trainIndex <- createDataPartition(ae$index, p = prop, 
                                  list = FALSE, 
                                  times = 1)
# assign partition
aeTrain <- ae[ trainIndex,]

# view training partition
aeTrain %>%
  # set the period column to show in Month-Year format
  mutate_at(vars(period), format, "%b-%y") %>% 
  # set the numeric columns to have a comma at the 1000's place
  mutate_at(vars(attendances, breaches), comma) %>%
  # show the first 10 rows
  head(10) %>%
  # format as a table
  kable()

# save training partition
write_csv(aeTrain, here("Data", "ae_attendances_WM_4hr_perfom_train.csv"))

# assign test partition
aeTest  <- ae[-trainIndex,]

# confirm length of test data
# nrow(aeTest)

# assign marker record
aeTestMarker  <- aeTest[1,]
# view
aeTestMarker  %>%
  # set the period column to show in Month-Year format
  mutate_at(vars(period), format, "%b-%y") %>% 
  # set the numeric columns to have a comma at the 1000's place
  mutate_at(vars(attendances, breaches), comma) %>%
  # show the first 10 rows
  head(10) %>%
  # format as a table
  kable()


# save marker record
write_csv(aeTestMarker, here("Data", "ae_attendances_WM_4hr_perfom_test_marker.csv"))

# reassign test partition without marker record
aeTest  <- aeTest[2:nrow(aeTest),]
# view
aeTest  %>%
  # set the period column to show in Month-Year format
  mutate_at(vars(period), format, "%b-%y") %>% 
  # set the numeric columns to have a comma at the 1000's place
  mutate_at(vars(attendances, breaches), comma) %>%
  # show the first 10 rows
  head(10) %>%
  # format as a table
  kable()

# save test data
write_csv(aeTest, here("Data", "ae_attendances_test.csv"))
```
# Data dictionary

### Read in data collected in Python
```{r read_data, echo =FALSE, message=FALSE}
# read data
CollectedData=read_csv(here("RawData", "CollectedDataFinal.csv"))
# view
glimpse(CollectedData)
```

### Build linker data frame with variable descriptions and types

``` {r variables_and_linker, echo=FALSE}
# variable descriptions
variable_description <- c("The index column that allows us to link the data collected to the original ae_attendances data in the 'RawData' folder.",
"The month that this activity relates to, stored as a date (1st of each month).",
"The Organisation data service (ODS) code for the organisation. If you want to know the organisation associated with a particular ODS code, you can look it up from the following address: https://odsportal.digital.nhs.uk/Organisation/Search.",
"The number of attendances for this department type at this organisation for this month.", 
"The number of attendances that breached the four-hour target.",  
"The number of attendances that resulted in an admission to the hospital.",
"The breach performance ([1 - breaches]/attendances)",
"The rate of admission (admissions/attendances)",
"The consent from the end-user to process and share the data collected with the data capture tool.")
print(variable_description)

# variable types
variable_type <- c(0, 1, 1, 0, 0, 0, 0, 0, 1)
print(variable_type)

# linker
linker<-build_linker(CollectedData, variable_description, variable_type)
print(linker)
```

### Use linker dataframe to create data dictionary

``` {r build_dictionary, echo=FALSE, results='hide'}
# build dictionary
dictionary <- build_dict(my.data = CollectedData, linker = linker)
```


``` {r view_save_dict, echo=FALSE}
# view

head(dictionary)
# save
write_csv(dictionary, here("RawData", "CollectedData_DataDictionary.csv"))
````

### Appending data dictionary to collected data

``` {r appending_data_dict, echo=FALSE}
# create main string to describe collected data
main_string <- "This data describes accident and emergency (A&E) metrics for hospital trusts within the chosen ambulance service region. Data source: NHSRdatasets package"
# view
main_string

# append dictionary
complete_CollectedData <- incorporate_attr(my.data = CollectedData, data.dictionary = dictionary,
main_string = main_string)
#Change the author name
attributes(complete_CollectedData)$author[1]<-"B209978"
# view data
complete_CollectedData
# view metadata
attributes(complete_CollectedData)
#save file
save_it(complete_CollectedData, here("RawData", "complete_CollectedData"))
#view file
complete_CollectedData<-readRDS(here("RawData", "complete_CollectedData.rds"))
```

# Data capture tool (Python)  

Perfomed in Python. Jupyter widgets designed to capture data from relevant hospitals. Performance and admission rate metrics calculated and inputted using simple block of Python code.  

Please find here: *https://github.com/B209978/B209978_assessment/tree/master/AssessmentFiles*