---
title: "Asessment Rmd file"
author: "B209978"
date: "17/06/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **R code to extract relevant data**
link to GitHub repo

## Loading packages including NHSR dataset
### Packages used are:

```{r Loading_datasets, message=FALSE, warning=FALSE}
# required packages are:
library(tidyverse)
library(NHSRdatasets)
library(here)
library(knitr)
library(scales)
library(caret)
library(dataMeta)
```

### Load and explore NHS England A&E attendance data
#### Examining structure and completeness of dataset
``` {r load_data, echo=FALSE}
# review dataset and display any missing values
data(ae_attendances)
ae<-ae_attendances
glimpse(ae)
ae
ae %>% 
  map(is.na) %>%
map(sum)
```

### Adding index for later linkage
``` {r index_and_save, echo=FALSE}
# add index column
ae <- rowid_to_column(ae, "index")
# display for confirmation
ae %>%
  # Set the period column to show in month-year format
  mutate_at(vars(period), format, "%b-%y") %>% 
  # Set the numeric columns to have a comma at the 1000's place
  mutate_at(vars(attendances, breaches, admissions), comma) %>%
  # Show the first 10 rows
  head(10) %>%
  # Format as a table
  kable()
# save file
write_csv(ae, here("RawData", "ae_attendances.csv"))
```

### Filter data
#### Removed all hospitals except consultant led Emergency Departments covered by selected ambulance service

``` {r, filter_data, echo=FALSE}
# filter to include certain hospitals
# filter to include type 1 hospitals
# filter to remove type column
ae<-ae %>% select(index, period, org_code, type, attendances, breaches, admissions)%>% 
  filter(org_code == "RKB"|org_code == "RRK"|org_code == "RXK"|org_code == "RLT"| org_code == "RJC"|
           org_code == "RWP"|org_code == "RLQ"|org_code == "RJE"| org_code == "RL4"|org_code == "RNA"|
           org_code == "RBK"|org_code == "RQW"|org_code == "RXW",type == 1) %>% 
  select(-type)

ae
```


### Calculate metrics

``` {r calculations, echo=FALSE}
# calculate and view breach performance
WM_performance <- ae %>%
  group_by(period) %>%
  summarise_at(vars(attendances, breaches), sum) %>%
  mutate(breach_performance = 1 - breaches / attendances)
glimpse(WM_performance)

# calculate and view admission rate
WM_admission <- ae %>%
  group_by(period) %>%
  summarise_at(vars(attendances, admissions), sum) %>%
  mutate(admission_rate = admissions / attendances)
glimpse(WM_admission)

# calculate and view regional attendances
WM_attendance <- ae %>%
  group_by(period) %>%
  summarise_at(vars(attendances), sum)
glimpse(WM_admission)
```

### Brief visualisation of regional patterns
``` {r graphs, echo=FALSE}
# plot performance
ggplot(WM_performance, aes(period, breach_performance)) +
  geom_line(color = "darkcyan") +
  geom_point(color = "darkcyan") +
  scale_y_continuous(labels = percent) +
  scale_x_date(date_labels = "%b-%y", date_breaks = "12 month")+
  labs(x = "Month of attendance",
       y = "% of A&E attendances that met the four hour standard",
       title = "Diminishing accident and emergency (A&E) performance",
       caption = "Source: NHSRdatasets")

# plot admission rate
ggplot(WM_admission, aes(period, admission_rate)) +
  geom_line(color = "darkcyan") +
  geom_point(color = "darkcyan") +
  scale_y_continuous(labels = percent) +
  scale_x_date(date_labels = "%b-%y", date_breaks = "12 month")+
  labs(x = "Month of attendance",
       y = "% of A&E attendances that result in admission",
       title = "Increasing accident and emergency (A&E) admission rate",
       caption = "Source: NHSRdatasets")

# plot regional admissions
ggplot(WM_attendance, aes(period, attendances)) +
  geom_line(color = "darkcyan") +
  geom_point(color = "darkcyan") +
  scale_y_continuous() +
  scale_x_date(date_labels = "%b-%y", date_breaks = "12 month")+
  labs(x = "Month of attendance",
       y = "Number of A&E attendances",
       title = "Increasing accident and emergency (A&E) attendances",
       caption = "Source: NHSRdatasets")
```

### Save data subset
``` {r save_subset, echo=FALSE}
# view subset
glimpse(ae)
# save subset
write_csv(ae, here("RawData", "ae_attendances_WM_4hr_perfom.csv"))
```

### Divide dataset for test, training and marker
``` {r divide_dataset, echo=FALSE}
# confirm length of dataset
# nrow(ae)


# calculate proportion needed for small training datset 
prop<-(1-(15/nrow(ae)))

# ensure consistant random number generator
set.seed(333)
# create training partition
trainIndex <- createDataPartition(ae$index, p = prop, 
                                  list = FALSE, 
                                  times = 1)
# assign partition
aeTrain <- ae[ trainIndex,]

# view training partition
aeTrain %>%
  # set the period column to show in Month-Year format
  mutate_at(vars(period), format, "%b-%y") %>% 
  # set the numeric columns to have a comma at the 1000's place
  mutate_at(vars(attendances, breaches), comma) %>%
  # show the first 10 rows
  head(10) %>%
  # format as a table
  kable()

# save training partition
write_csv(aeTrain, here("Data", "ae_attendances_WM_4hr_perfom_train.csv"))

# assign test partition
aeTest  <- ae[-trainIndex,]

# confirm length of test data
# nrow(aeTest)

# assign marker record
aeTestMarker  <- aeTest[1,]
# view
aeTestMarker  %>%
  # set the period column to show in Month-Year format
  mutate_at(vars(period), format, "%b-%y") %>% 
  # set the numeric columns to have a comma at the 1000's place
  mutate_at(vars(attendances, breaches), comma) %>%
  # show the first 10 rows
  head(10) %>%
  # format as a table
  kable()


# save marker record
write_csv(aeTestMarker, here("Data", "ae_attendances_WM_4hr_perfom_test_marker.csv"))

# reassign test partition without marker record
aeTest  <- aeTest[2:nrow(aeTest),]
# view
aeTest  %>%
  # set the period column to show in Month-Year format
  mutate_at(vars(period), format, "%b-%y") %>% 
  # set the numeric columns to have a comma at the 1000's place
  mutate_at(vars(attendances, breaches), comma) %>%
  # show the first 10 rows
  head(10) %>%
  # format as a table
  kable()

# save test data
write_csv(aeTest, here("Data", "ae_attendances_test.csv"))
```
# Data dictionary


# Data capture tool (Python)  

Perfomed in Python. Jupyter widgets designed to capture data from relevant hospitals. Performance and admission rate metrics calculated and inputted using simple block of Python code.